Sensitivity Analysis

### 2024-04-22 SFE_article_Sensitivity_Analysis_short - Chamomile/Sections/Results_Chamomile.tex
22.04

	In both scenarios previously discussed, the fitting outcomes were not deemed satisfactory. Building upon the concepts underlying the Broken-and-Intact and Shrinking Core models (detailed in Chapter \ref{CH: Gamma_Function}), the gamma function is introduced to capture the decreasing extraction kinetics. The correction factor is combined with the simplified linear model, resulting in a two-parameter model ($D_i^R$ and $\Upsilon$) as given by Equation \ref{Model_kinetic}. Figure \ref{fig: Fit_1_Di_Gamma} shows the parameter space and the corresponding cost function values.

@@

	\citet{Rahimi2011} analysed the same dataset using the desorption–dissolution–diffusion model. To decrease the number of unknown parameters, \citet{Rahimi2011} applied a set of empirical correlations. The remaining parameters were determined through the use of a genetic algorithm to solve the parameter estimation problem for each experiment individually. Results obtained by \citet{Rahimi2011} shows that the desorption–dissolution–diffusion model fails to reproduce the yield data. The main difference between the model employed by \citet{Rahimi2011} and the one in this study is the $\gamma$ function. The $\gamma$ function increases the model's flexibility by adjusting $D_i$ and allows it to obtain a better fit.

>>>>

	In both scenarios previously discussed, the fitting outcomes were not deemed satisfactory. Building upon the concepts underlying the Broken-and-Intact and Shrinking Core models (detailed in Chapter \ref{CH: Gamma_Function}), the $\gamma$-function is introduced to capture the decreasing extraction kinetics. The correction factor is combined with the simplified linear model, resulting in a two-parameter model ($D_i^R$ and $\Upsilon$) as given by Equation \ref{Model_kinetic}. Figure \ref{fig: Fit_1_Di_Gamma} shows the parameter space and the corresponding cost function values.

@@

	\citet{Rahimi2011} analysed the same dataset using the desorption–dissolution–diffusion model. To decrease the number of unknown parameters, \citet{Rahimi2011} applied a set of empirical correlations. The remaining parameters were determined through the use of a genetic algorithm to solve the parameter estimation problem for each experiment individually. Results obtained by \citet{Rahimi2011} shows that the desorption–dissolution–diffusion model fails to reproduce the yield data. The main difference between the model employed by \citet{Rahimi2011} and the one in this study is the $\gamma$-function. The $\gamma$-function increases the model's flexibility by adjusting $D_i$ and allows it to obtain a better fit.

### 2024-04-22 SFE_article_Sensitivity_Analysis_short - Chamomile/Sections/Results_Sensitivity.tex
22.04

   Figure \ref{fig:Sensitivty_F_CS} visualizes the local sensitivity of the solute concentration in the solid phase to changes in the mass flow rate during a supercritical fluid extraction process. It is postulated that with an increase in the mass flow rate, the driving force for mass transfer—the concentration gradient—becomes more pronounced, leading to an accelerated decrease in solute concentration within the solid matrix. This is exhibited by the negative sensitivity values: the solute is depleted more rapidly from the solid phase with higher mass flow rates. The most significant sensitivity is seen at the process commencement, particularly at the column inlet (top left corner with bright colours), where the concentration of solute is highest, and the effect of mass flow rate on extraction is most pronounced.

   As the extraction progresses, the sensitivity values become less negative, moving towards zero, depicted by the gradient transitioning to cooler colors towards the bottom of the figure. This gradual change corresponds to the decreasing concentration of the solute in the solid phase, which results in a diminishing concentration gradient and, consequently, a reduction in the rate of extraction. The sensitivity approaching zero suggests that the driving force for mass transfer is decreasing as the solute is being depleted.

   The asymptotic approach to zero sensitivity indicates a physical plateau in the extraction process, where further increases in mass flow rate no longer enhance the extraction efficiency. In practical terms, this asymptotic behavior signifies the approach towards an extraction endpoint where the remaining solute concentration in the solid phase is minimal and becomes less dependent on the mass flow rate. At this stage, other factors, such as diffusion within the solid matrix or equilibrium limitations, may limit the extraction rate.

   In the most extreme cases, depicted by the uniform regions in the plot, the solute has been effectively exhausted, and the remaining traces in the solid phase are no longer influenced by changes in the mass flow rate. This stage represents the final equilibrium, where the concentration gradient effectively drops to zero, rendering the mass flow rate inconsequential for the extraction of the remaining solute. At this juncture, the system has reached its thermodynamic limit, and the efficiency of extraction cannot be enhanced by merely increasing the flow rate. This underscores the importance of optimizing the mass flow rate at the beginning and throughout the extraction process to ensure efficient solute recovery before reaching this terminal stage.

@@

    The Figure \ref{fig:Sensitivty_F_CF} depicts the sensitivity of the solute concentration in the fluid phase to the flow rate within a supercritical fluid extraction system. This sensitivity analysis reveals how the solute's behavior in the fluid phase is influenced by variations in the flow rate over the course of the extraction process, as depicted by both time and the length of the extraction column.

    In the initial phase of the extraction, the sensitivities near the inlet are negligible, as shown by the colors being close to zero. This indicates a minimal initial response to changes in the flow rate, which could be due to the presence of an high concentration gradient, allowing the fluid phase to rapidly reach solute regardless of the flow rate.

    As the extraction progresses, positive sensitivities emerge along the length of the extractor, indicated by the concentration front that develops and moves along the flow direction. This front represents the increased movement of the solute with the fluid phase as the flow rate increases, suggesting an enhanced mass transfer due to the elevated velocity of the supercritical fluid. This velocity increase augments the concentration gradient, thus encouraging a more efficient solute migration from the solid to the fluid phase.

    Following the positive front, there is an emergence of a front characterized by negative sensitivities. This trend illustrates a diminishing solute concentration in the fluid phase, and it reflects a point in the extraction where the solute availability in the solid phase becomes a limiting factor; the depletion of solute reduces the concentration gradient, thereby decelerating the extraction rate.

>>>>

   Figure \ref{fig:Sensitivty_F_CS} visualizes how the solute concentration in the solid phase is affected by the mass flow change. It is postulated that with an increase in the mass flow rate, the concentration gradient becomes larger leading to an accelerated decrease of the solute from solid particles and it is characterized by negative sensitivity values. Gradually, the sensitivities decrease until they reach minimum values and the effect of mass flow rate on extraction is most significant.

   As the extraction progresses, the sensitivity values become less negative, moving towards zero. This gradual change corresponds to the decreasing concentration of the solute in the solid phase, which results in a diminishing concentration gradient and, consequently, a reduction in the rate of extraction. The sensitivity approaching zero suggests that the driving force for mass transfer is decreasing as the solute is being depleted. The asymptotic behaviour can be interpreted as a point where the remaining solute concentration in the solid phase is minimal and becomes less dependent on the mass flow rate. At this stage, other factors, such as internal diffusion within the solid matrix or equilibrium limitations, may limit the extraction rate.

   %The asymptotic approach to zero sensitivity indicates a physical plateau in the extraction process, where further increases in mass flow rate no longer enhance the extraction efficiency. In practical terms, this asymptotic behavior signifies the approach towards an extraction endpoint where the remaining solute concentration in the solid phase is minimal and becomes less dependent on the mass flow rate. At this stage, other factors, such as diffusion within the solid matrix or equilibrium limitations, may limit the extraction rate.

   %In the most extreme cases, depicted by the uniform regions in the plot, the solute has been effectively exhausted, and the remaining traces in the solid phase are no longer influenced by changes in the mass flow rate. This stage represents the final equilibrium, where the concentration gradient effectively drops to zero, rendering the mass flow rate inconsequential for the extraction of the remaining solute. At this juncture, the system has reached its thermodynamic limit, and the efficiency of extraction cannot be enhanced by merely increasing the flow rate. This underscores the importance of optimizing the mass flow rate at the beginning and throughout the extraction process to ensure efficient solute recovery before reaching this terminal stage.

@@

    The Figure \ref{fig:Sensitivty_F_CF} depicts the sensitivity of the solute concentration in the fluid phase to the flow rate within a supercritical fluid extraction system. In the initial phase of the extraction, the sensitivities near the inlet are negligible, as characterized by values close to zero. This indicates a minimal initial response to changes in the flow rate, which could be due to the presence of an high concentration gradient, allowing the fluid phase to rapidly reach solute regardless of the flow rate.

    As the extraction progresses, positive sensitivities emerge along the length of the extractor and form the concentration front that moves in the flow direction. This front represents the increased movement of the solute with the fluid phase as the flow rate increases, suggesting an enhanced mass transfer due to the elevated velocity of the supercritical fluid. Following the positive front, there is an emergence of a front characterized by negative sensitivities. This trend illustrates a diminishing solute concentration in the fluid phase, and it reflects a point in the extraction where the solute availability in the solid phase becomes a limiting factor; the depletion of solute reduces the concentration gradient, thereby decelerating the extraction rate. \todo{here}

### 2024-04-22 SFE_article_Sensitivity_Analysis_short - Chamomile/Sections/introduction_imp.tex
22.04

	This study investigates the extraction of essential oil from chamomile flowers (Matricaria chamomilla L.) via supercritical fluid extraction techniques and the modelling of this process. Chamomile is a medicinal herb widely cultivated in southern and eastern Europe—such as Germany, Hungary, France, and Russia. It can be found outside of Europe in Brazil as discussed by \citet{Singh2011}. This plant is distinguished by its hollow, bright gold cones, housing disc or tubular florets and surrounded by about fifteen white ray or ligulate florets. Chamomile has been used for its medicinal benefits, serving as an anti-inflammatory, antioxidant, mild astringent, and healing remedy. Chamomile's aqueous extract is widely used as a gentle sedative to calm nerves and mitigate anxiety, hysteria, nightmares, insomnia, and other sleep-related conditions, according to \citet{Srivastava2009}. \citet{Orav2010} reported that oil yields from dried chamomile samples ranged from 0.7 to 6.7 mL/kg. Notably, the highest yields of essential oil, between 6.1 and 6.7 mL/kg, were derived from chamomile sourced from Latvia and Ukraine, while chamomile from Armenia exhibited a lower oil content of 0.7 mL/kg.

@@

	Evaluating the economic viability of the process is essential when choosing the suitable technology for essential oil extraction. Traditional methods, such as distillation and organic solvent extraction, are commonly employed but come with drawbacks. Distillation, for example, involves high temperatures that can lead to the thermal degradation of heat-sensitive compounds. This limitation has led to the increased popularity of alternative techniques like supercritical fluid extraction. Supercritical carbon dioxide is particularly appealing due to its distinctive properties: it is inflammable, non-toxic, and is non-corrosive. Moreover, compared to other fluids, its relatively low critical point positions it as an advantageous alternative to conventional extraction methods. Supercritical fluids are capable of exhibiting both gas- and liquid-like properties, allowing for adjustable dissolving power through changes in operating conditions.

	The literature offers various mathematical models to detail the extraction of valuable compounds from a fixed biomass bed. Selecting a fitting model necessitates a deep understanding of the physical processes at play within the operational unit, as each model is built on specific assumptions and outlines distinct mass transfer mechanisms and equilibrium dynamics.

	One model proposed by \citet{Reverchon1993} is the hot ball model, which is based on an analogy to heat transfer and describes an extraction process from solid particles containing small quantities of solute where solubility is not a limiting factor.

@@

	This study focuses on creating a process model for the extraction of natural substances from solid materials using supercritical fluids, with a particular emphasis on supercritical CO$_2$. The approach involves estimating the solvent properties through thermodynamic relationships and determining the extraction kinetic parameters via a series of experiments conducted under a variety of conditions. The maximum likelihood estimator method is employed to address the parameter estimation challenge, with the derived parameters subsequently used to establish correlations. These correlations are instrumental in extending the applicability of the process model across different temperatures (30 to 40$~^\circ \text{C}$), pressures (100 to 200 bar) and flow rates (3.33 to 6.67 $\times 10^{-5}$ kg/s).

>>>>

	This study investigates the extraction of essential oil from chamomile flowers (Matricaria chamomilla L.) via supercritical fluid extraction techniques and the modelling of this process. Chamomile is a medicinal herb widely cultivated in southern and eastern Europe—such as Germany, Hungary, France, and Russia. It can be found outside of Europe in Brazil as discussed by \citet{Singh2011}. This plant is distinguished by its hollow, bright gold cones, housing disc or tubular florets and surrounded by about fifteen white ray or ligulate florets. Chamomile has been used for its medicinal benefits, serving as an anti-inflammatory, antioxidant, mild astringent, and healing remedy. Chamomile's aqueous extract is widely used to calm nerves and mitigate anxiety, hysteria, nightmares, insomnia, and other sleep-related conditions, according to \citet{Srivastava2009}. \citet{Orav2010} reported that oil yields from dried chamomile samples ranged from 0.7 to 6.7 mL/kg. The highest yields of essential oil, between 6.1 and 6.7 mL/kg, were derived from chamomile sourced from Latvia and Ukraine, while chamomile from Armenia exhibited a lower oil content of 0.7 mL/kg.

@@

	Evaluating the economic viability of the process is essential when choosing the suitable technology for essential oil extraction. Traditional methods, such as distillation and organic solvent extraction, are commonly employed but come with drawbacks. Distillation, for example, involves high temperatures that can lead to the thermal degradation of heat-sensitive compounds. This limitation has led to the increased popularity of alternative techniques like supercritical fluid extraction. Supercritical carbon dioxide is appealing due to its distinctive properties: it is inflammable, non-toxic, and is non-corrosive.  CO$_2$ is the most used supercritical fluid, sometimes modified by co-solvents such as ethanol. Supercritical fluids are capable of exhibiting both gas- and liquid-like properties, allowing for adjustable dissolving power through changes in operating conditions.

	The literature offers various mathematical models to describe the extraction of valuable compounds from a biomass. Selecting a model requires a deep understanding of the physical processes, as each model is built on specific assumptions, mass transfer mechanisms and equilibrium dynamics.

	The model proposed by \citet{Reverchon1993} is the hot ball model, which is based on an analogy to heat transfer and describes an extraction process from solid particles. The model is based on assumptions that that particles contains low quantities of solute and solubility is not a limiting factor.

@@

	This study focuses on finding a process model for the extraction of natural substances from solid materials using supercritical fluids, with a particular emphasis on supercritical CO$_2$. The approach involves estimating the solvent properties through thermodynamic relationships and determining the extraction kinetic parameters via a series of experiments conducted under a variety of conditions. The maximum likelihood estimation method is employed solve the parameter estimation problem. Later, the correlations between parameters and operating conditions are found.
